---
title: "Working with the Exponential Power Distribution Using `gnorm`"
author: "Maryclare Griffin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with the Exponential Power Distribution Using `gnorm`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The exponential power distribution, also known as the [generalized normal distribution](https://en.wikipedia.org/wiki/Generalized_normal_distribution), was first described a very long time ago in Subbotin (1923)^[[M. T. Subbotin. "On the Law of Frequency of Error." Mat. Sb. 31.2 (1923), 206-301.](http://www.mathnet.ru/php/archive.phtml?wshow=paper&jrnid=sm&paperid=6854&option_lang=eng)] and rediscovered as the generalized normal distribution in Nadarajah (2005)^[[Nadarajah, Saralees. "A generalized normal distribution." Journal of Applied Statistics 32.7 (2005):685-694.](http://www.tandfonline.com/doi/abs/10.1080/02664760500079464)]. It generalizes the Laplace, normal and uniform distributions and is pretty easy to work with in many ways, so it can be very useful. Accordingly, I've made a little `R` package called `gnorm` that provides density, CDF and quantile functions for the exponential power distribution as well as random variate generation. We can load the package from GitHub as follows:
```{r, message = FALSE}
library(devtools)
install_github("maryclare/gnorm")
library(gnorm)
```

Since we're going to be generating some random variables, we should also set a seed:
```{r}
set.seed(100)
```

# Exponential Power Density

An exponential power distributed random variable, $x$, has the following density:

$$
p(x | \mu, \alpha, \beta) = \frac{\beta}{2\alpha \Gamma(1/\beta)}\text{exp}\{-(\frac{x - \mu}{\alpha})^\beta\},
$$
where $\mu$ is the mean of $x$ and $\alpha > 0$ and $\beta > 0$ are scale and shape parameters.

The function `dgnorm` gives the exponential power density evaluated at specified values of $x$, $\mu$, $\alpha$ and $\beta$.

```{r}
dgnorm(x = 1, mu = 0, alpha = 1, beta = 1)
```


An aside - this is not my preferred way of parametrizing the exponential power density. I prefer to replace $\alpha$ and $\beta$ with the standard deviation of $x$, $\sigma$, and a shape parameter $q$:
$$
p(x | \mu, \sigma, q) = \frac{q}{2\tau}\sqrt{\frac{\Gamma(3/q)}{\Gamma(1/q)^3}}\text{exp}\{-(\frac{\Gamma(3/q)}{\Gamma(1/q)})^{q/2}(\frac{x - \mu}{\sigma})^q\}
$$

This is similar but not identically to the parametrization preferred by Box and Tiao (1973)^[[G. E. P. Box and G. C. Tiao. "Bayesian inference in Statistical Analysis." Addison-Wesley Pub. Co., Reading, Mass (1973).](http://onlinelibrary.wiley.com/book/10.1002/9781118033197)]. That said, the parametrization using $\alpha$ and $\beta$ is the one featured on Wikipedia so realistically, it's the one everyone will find when they look up this distribution, so I'll use the $\alpha$ and $\beta$ parametrization for the rest of the vignette.
